The Werewolf Among Us: Humans vs LLMs in Multi-Agent Games

### 2025-05-06 Quarto/paper.qmd
Methods

- We wanted to compare how humans and large language models (LLMs) handle these situations. To do this, we used two recent datasets: 
	- _Werewolf Among Us_ (Lai et al., 2022), a collection of real human gameplay annotated with persuasion strategies, 
	- _Werewolf Arena_ (Bailis et al., 2024), a simulated environment where LLM agents play the game autonomously. 
	- While both studies show Werewolf generates complex strategic language, neither compares human and LLM behavior directly.

@@

- Human dataset description [@laiWerewolfUsMultimodal2022]
- Is specifically for a form of one-night werewolf
  - Describe key differences
- Used specifically for the text available
  - and annotations of persuasion strategy on the text

### Werewolf Arena

- [@bailisWerewolfArenaCase2024]
- Discuss the framework, how it works, prompts, etc
- Discuss what types of runs we did
- Discuss the data included in output
- Talk about how we had to annotate the LLM speech with persuasion strategies ourselves

@@

- Formatted data to match, performed various comparisons

>>>>

- We wanted to compare how humans and large language models (LLMs) handle these situations. To do this, we used a recent human dataset and constructed one from an LLM-based simulator: 
	- _Werewolf Among Us_ (Lai et al., 2022), a collection of real human gameplay annotated with persuasion strategies, 
	- _Werewolf Arena_ (Bailis et al., 2024), a simulated environment where LLM agents play the game autonomously. 
	- While both studies show Werewolf generates complex strategic language, neither compares human and LLM behavior directly.

@@

We used the _Werewolf Among Us_ dataset (Lai et al., 2022), which contains annotated dialogues from over 150 real games of One Night Werewolf and Avalon. These games differ from classic Werewolf because they have only one round of discussion and voting, do not eliminate players during the game, and include many specialized roles beyond Villager and Werewolf. The dataset provides detailed annotations of persuasion strategies for each utterance, such as accusations, defenses, and identity claims. We specifically used the textual transcriptions and their strategy annotations for our comparisons.

### Werewolf Arena (LLM Dataset)

The _Werewolf Arena_ dataset (Bailis et al., 2024) features simulated classic Werewolf games played by LLM agents without human intervention. Unlike the one-round human games, these simulations involve multiple rounds alternating between night (secret actions) and day (open discussion). Each LLM agent is assigned a role (Villager, Werewolf, Seer, Doctor) and receives context-specific prompts that guide their strategic interactions.

We ran simulations using five LLM models: GPT-4o, GPT-4.1, GPT-4o-mini, DeepSeek-Chat, and DeepSeek-Reasoner. We experimented with two configurations:

- 8 players with 8 rounds of discussion

- 10 players with 6 rounds of discussion

We chose these configurations to give villagers more time to coordinate, extending the dialogue to better observe persuasive behaviors. After running the simulations, we manually annotated the LLM-generated dialogue using the same persuasion categories as in the human dataset.

@@

We standardized annotations across both datasets so they could be directly compared. Our analyses focused on frequency distributions of persuasion strategies, comparisons by role (villager versus werewolf), and differences in how humans and LLMs applied these strategies.
